---
title: "R Notebook"
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---

---
title: "hw8"
output: html_document
---

## Question 11.1

Using the crime data set uscrime.txtfrom Questions 8.2, 9.1, and 10.1, build a regression model using

#### Stepwise regression

```{r}
setwd("c:\\stuff")
uscrime=read.table("uscrime.txt",header=T)

current=names(uscrime)[-16]
for (i in 1:12) {
  form=as.formula(paste("Crime ~ ",paste(current, collapse="+")))
  #print(form)
  reg=lm(form,data=uscrime)
  prt=summary(reg)$coefficients[,'Pr(>|t|)']
  print(prt)
  max=0.05
  new_i=-1
  for (i in 2:length(prt)) {
   print(paste(prt[i],max))
   if (prt[i] > max) {
      max=prt[i]
      new_i = i
    }
  }
  if (new_i != -1) {
    #remove the factor
    print("removing factor")
    print(new_i)
    print(paste(paste(prt[new_i],""),current[(new_i-1)]))
    current=current[-(new_i-1)]
  }
}
print("FINAL REGRESSION")
summary(reg)
```


## Question 11.1 (cont.)

Using the crime data set uscrime.txtfrom Questions 8.2, 9.1, and 10.1, build a regression model using:1.Stepwise regression2.Lasso3.Elastic netFor Parts 2 and 3, remember to scale the data first –otherwise, the regression coefficients will be on different scales andthe constraint won’t have the desired effect.

Reference::: https://web.stanford.edu/~hastie/Papers/Glmnet_Vignette.pdf

We do LASSO first this is alpha=1

If you use glmnet, the scaling is performed by the package. You don't need to worry about scaling the test set because the "coefficients are always returned on the original scale".

For this not doing a training and test set

```{r}
rm(list=ls())
set.seed(1)
setwd("c:\\stuff")
uscrime=read.table("uscrime.txt",header=T)

#uscrime=scale(uscrime,center=FALSE)
indep=as.matrix(uscrime[,1:15])
depend=as.matrix(uscrime[,16])
library(glmnet)
fit=glmnet(indep,depend,family="gaussian",alpha=1)

#run cross validation to choose a model
cvfit=cv.glmnet(indep,depend,family="gaussian",alpha=1)
#plot the values and see the selected lambdas
plot(cvfit)

#get the coefficients for the cross validated model
coef(cvfit, s = "lambda.min")


```

Next we do Elastic net with an alpha of 0.2, 0.5, 0.7

```{r}
cvfit=cv.glmnet(indep,depend,family="gaussian",alpha=.5)
plot(cvfit)
cvfit=cv.glmnet(indep,depend,family="gaussian",alpha=.7)
plot(cvfit)
cvfit=cv.glmnet(indep,depend,family="gaussian",alpha=.2)
plot(cvfit)
coef(cvfit, s = "lambda.min")
```

0.2 looks the best with the mean squared error around the lambda


```{r}
coef(cvfit, s = "lambda.min")
```

