---
title: "Assignment Three"
runtime: shiny
output: pdf_document
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
```
## Question 5.1

### Grubbs.test to find outliers

#### Plot the original data to see the outlier

```{r}
UScrime <- read.delim("http://www.statsci.org/data/general/uscrime.txt")
crime_info=UScrime[,'Crime']
crime_vector=as.integer(unlist(crime_info[1]))
library('outliers')
grubbs.test(crime_info,10)
summary(crime_info)
plot(crime_info)
```

### Elbow Graph for optimal number of clusters
#### Great reference https://uc-r.github.io/kmeans_clustering
#### Deep analysis for optimal clusters \
#### https://www.r-bloggers.com/finding-optimal-number-of-clusters/

```{r}
library(purrr)
set.seed(123)

# function to compute total within-cluster sum of square 
k_means=function(k) {
  kmeans(crime_info, k, nstart = 10 )$tot.withinss
}

range_of_k <- 1:20

withinss <- map(range_of_k, k_means)

plot(range_of_k, withinss,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
```

## Show clusters
#### print clusters https://www.r-bloggers.com/k-means-clustering-in-r/

```{r}
library(fpc)
km_final=kmeans(crime_info, 5, nstart = 10)
km_final$cluster <- as.factor(km_final$cluster)
plotcluster(crime_info,km_final$cluster)
```

## Question 6.1

I am a very average poker player but like most would love to get better.  I do not have unlimted bugdet but like to challenge myself at the highest level beyond my talents.  As I can not go broke but need experience so I need to monitor my budget (stack) which has to be of a bigger size that I am willing to lose to get into higher valued games.  Further there is randomness that effects my staff even if I am playing well as people get lucky with cards.  CUSUM can be used to set a mean and a tolerance so that if I am playing well of poorly and losing on bad beats I can stop for the day and live to fight another day.

Take the mean to be the loss of my skill level on a full level 10 hands (say this is 2 times what is called the big blind).  Take the threshold to be twice the mean.  If I break the threshold I am done for the day.

## Question 6.2
### reference https://itl.nist.gov/div898/handbook/pmc/section3/pmc323.htm
### good reference https://www.measurementlab.net/publications/CUSUMAnomalyDetection.pdf

1.  Answer from the average of each day over all years is day 112 which is 112-62-30=20

October 20th this seems late

```{r}
a=read.csv("1862052.csv")
m=matrix(unlist(strsplit(as.character(a$DATE),'-')), ncol=3,byrow=TRUE)
b=data.frame(m)
b$TMAX=a$TMAX
y=b[grep("07|08|09|10", b$X2),]
colnames(y)<-c("Year","Month","Day","Temp")
y=y[!grepl('2016',y$Year),]


#qcc not used cusum(data=y$Temp, center=mean(y$Temp), std.dev=sd(y$Temp), decision.interval = 5)

library(tidyverse)
hu=y %>% group_by(Month,Day) %>% summarize(mean(Temp)) %>% data.frame
colnames(hu)<-c('Month','Day','Temp')
mu=mean(hu$Temp)
k=sd(hu$Temp)
T=4*sd(y$Temp)
S=0;count=0; for (i in hu$Temp) { S=max(0,S+(mu-i-k)); count=count+1 ;if (S > T) { print(S); print(count); break;} }
```


2.  Use a CUSUM approach to make a judgment of whether Atlanta’s summer climate has gotten
warmer in that time (and if so, when).

I think my conclusion is wrong as it does not say that Atlanta is warming.
With T equal k it does though in 2012
```{r}
library(tidyverse)
a=read.csv("1862052.csv")
m=matrix(unlist(strsplit(as.character(a$DATE),'-')), ncol=3,byrow=TRUE)
b=data.frame(m)
b$TMAX=a$TMAX
y=b
colnames(y)<-c("Year","Month","Day","Temp")
y=y[!grepl('2016',y$Year),]
hu=y %>% group_by(Year) %>% summarize(mean(Temp)) %>% data.frame
colnames(hu)<-c('Year','Temp')
mu=mean(hu$Temp)
k=sd(hu$Temp)
T=4*k
S=0;count=0; for (i in hu$Temp) { S=max(0,S+(i-mu-k)); count=count+1 ;if (S > T) { print(S); print(count); break;} }
hu
```


